---
title: "A4_R_Code"
author: "Advait Shah"
date: "12/02/2023"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r}

#a = read.csv(file.choose())

a = read.csv("Case5-data.csv")
```

```{r}
head(a)

```
```{r}
days=c("Mon", "Tue", "Wed", "Thu", "Fri")
for (i in 1:5){
    df2=a[a[,2]==days[i],]
    print(hist(df2[,"Actual"],main = paste("Histogram of" , days[i])))
}

```



```{r}
#make this example reproducible
set.seed(1)

#use 80% of dataset as training set and 20% as test set
sample <- sample(c(TRUE, FALSE), nrow(a), replace=TRUE, prob=c(0.8,0.2))
train  <- a[sample, ]
test   <- a[!sample, ]

t5_model = lm(Actual~T...5, train)
print(paste("T-5 based model R-Squared:",summary(t5_model)$r.squared))
print(paste("T-5 based model MSE:",mean((test$Actual - predict.lm(t5_model, test)) ^ 2)))
plot(train$T...5,train$Actual)
abline(t5_model, col = "blue")

t6_model = lm(Actual~T...6, train)
print(paste("T-6 based model R-Squared:",summary(t6_model)$r.squared))
print(paste("T-6 based model MSE:",mean((test$Actual - predict.lm(t6_model, test)) ^ 2)))
plot(train$T...6,train$Actual)
abline(t6_model, col = "blue")

t7_model = lm(Actual~T...7, train)
print(paste("T-7 based model R-Squared:",summary(t7_model)$r.squared))
print(paste("T-7 based model MSE:",mean((test$Actual - predict.lm(t7_model, test)) ^ 2)))
plot(train$T...7,train$Actual)
abline(t7_model, col = "blue")

t8_model = lm(Actual~T...8, train)
print(paste("T-8 based model R-Squared:",summary(t8_model)$r.squared))
print(paste("T-8 based model MSE:",mean((test$Actual - predict.lm(t8_model, test)) ^ 2)))
plot(train$T...8,train$Actual)
abline(t8_model, col = "blue")

t9_model = lm(Actual~T...9, train)
print(paste("T-9 based model R-Squared:",summary(t9_model)$r.squared))
print(paste("T-9 based model MSE:",mean((test$Actual - predict.lm(t9_model, test)) ^ 2)))
plot(train$T...9,train$Actual)
abline(t9_model, col = "blue")

```
[1] "T-5 based model R-Squared: 0.827860438875018"
[1] "T-5 based model MSE: 41.6998887704155"
[1] "T-6 based model R-Squared: 0.819831417075027"
[1] "T-6 based model MSE: 40.7105698327936"
[1] "T-7 based model R-Squared: 0.81812575168335"
[1] "T-7 based model MSE: 47.9691965433956"
[1] "T-8 based model R-Squared: 0.808648102575112"
[1] "T-8 based model MSE: 58.2459247847008"
[1] "T-9 based model R-Squared: 0.785580610181517"
[1] "T-9 based model MSE: 66.1200884449541"

As we can see, T-5 and T-6 gives the least mean square error in predicting the actual surgeries when tested on the test dataset. whereas, T-7, T-8, T-9 predictors based model gave comparatively higher prediction error. So, we could argue that it is a trade off between accuracy level and prediction timing.
We also plotted the best fit linear regression models based on each of these predictors, and calculated R-Squared values. It also suggests that linear regression model could be best fit on the T-5 and T-6 based predictors and thus, they gave better R-square values compared to T-7, T-8, T-9 predictors. This also explains why we have better prediction accuracy with T-5 and T-6 based models, as their linear regression models are more robust due to less bias in training dataset.

I would suggest T-6 based model to be used for predictions, because it has high prediction accuracy as well as team will have this information 6 days in advance of surgery, which would be sufficient in planning the schedule accurately about 6 days in advance of actual surgery days.


```{r}

# checking correlation between raw predictors
cor(a[,3:19])

```

This suggests predictors are highly correlated

```{r}


#transforming columns to reduce correlation and create new predictors
df = data.frame(a$DOW,a$Actual)
for (i in 3:18){
  df[,i] <- a[,i+1] - a[,i]
}

head(df)

```

```{r}

cor(df[,2:18])

```
the predictors are not correlated now, and hence we will include all of these predictors in our actual surgeries linear regression prediction.

```{r}
#removing DOW from df for model 1

df1 = df[c(-1)]
head(df1)

model_1_not_str = lm(a.Actual~., df1)

summary(model_1_not_str)

```


So, we are getting 74.93 R-squared value in this linear regression fit.
But, this model would not be much helpful in predicting surgeries in advance, because its predictors require data of daily addition in scheduled surgeries even close to surgery day.


```{r}

# model 2: Stratified by the day of the week

DOW = c("Mon","Tue","Wed","Thu","Fri")

for (i in 1:5){
  df2 = df[df[,1]==DOW[i],]
  df2 = df2[c(-1)]
  model1 = lm(a.Actual~., df2)
  print(paste(DOW[i],"Multiple LR model R-squared value:",summary(model1)$r.squared))
  
}


```

Other Possible Models as discussed in the lecture: 

Model 1 : Does not stratify by day of the week and use just T-7 as predictor

```{r}
t7_model = lm(Actual~T...7, a)
print(paste("T-7 based LR model R-squared value:",summary(t7_model)$r.squared))

```

Model 2 : Includes day of the week as dummy variables


```{r}
t7d_model = lm(Actual~T...7+DOW, a)
print(paste("T-7+DOW based LR model R-squared value:",summary(t7d_model)$r.squared))

```

Model 3 : stratify by day of the week and use just T-7 as predictor

```{r}
DOW = c("Mon","Tue","Wed","Thu","Fri")

for (i in 1:5){
  df3 = a[a[,2]==DOW[i],]
  t7s_model = lm(Actual~T...7, df3)
  print(paste(DOW[i],"stratified T-7 based LR model R-squared value:",summary(t7s_model)$r.squared))
  
}
```

################################

Now, consider the surgery (actual) date as a time series with September 4th to September 14th as the testing set and the rest as the training set. Fit a Moving Average (MA) model to the time series and visualize it.

```{r}
library(dplyr)
library(magrittr)
a %<>%
  mutate(SurgDate= as.Date(SurgDate, format= "%d-%m-%Y"))
```

```{r}

train <- subset(a, SurgDate < "2012-09-04")
test <- subset(a, SurgDate >= "2012-09-04")

tail(train)

```


```{r}
library(zoo)


s <- a %>%
  select(SurgDate, srate = Actual) %>%
  mutate(srate_tma = rollmean(srate, k = 3, fill = NA, align = "right"))


```

```{r}
library(ggplot2)

ggplot(s, aes(SurgDate,srate)) +
  geom_line() +
  theme_minimal()

```

```{r}
# let's use T-6 based model, which had given us least MSE of prediction on test data
summary(t6_model)

print(paste("T-6 based model MSE:",mean((test$Actual - predict.lm(t6_model, test)) ^ 2)))

```



